SunPy - Python for Solar Physics

### 2014-01-28 1-Introduction.tex

\section{Introduction}
Science, including solar physics and astrophysics, is driven by the analysis 
and summarizing of data. Modern advances in sensor technology combined with 
the availability of cheap storage and large bandwidth communications has led to 
a rapid and overwhelming increase in amount of data which scientists are responsible
for analyzing. For example, NASA's Solar 
Dynamics Observatory (SDO) satellite records over 1 TB of data per day. To analyze this 
mountain of data and enable scientific discovery requires increasingly complex software 
tools. Standardised, easy to use,
and transparent software tools are a key enabling technology in the scientific endeavor
so that the community can build upon a common foundation.
%schriste - the opening paragraph still needs work

SunPy aims to provide a free, open source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breath of high-quality scientific focused packages 
available for it. The Python programming language is a general purpose, 
powerful and easy to learn high-level programming language. Python is one of 
the top ten most popular programming languages in the world according to the 
2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
and is widely used outside of scientific fields in web development, education 
and `big data' analytics.

The development of a package such as SunPy in Python is made possible by the 
rich scientific Python ecosystem of packages available. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language, 
\textit{i.e.} array manipulation, core numerical algorithms and visualisation. 
Build upon these foundations, packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

The design philosophy of SunPy is to provide a clean, simple to use, and well 
structured package that provides the \textit{core} tools for solar physics. The 
primary focus of SunPy's early development is to provide specialised, linked, 
datatypes that allow the acquisition, processing and visualisation of all types 
of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project as well as future plans. The latest release of SunPy (0.4) from
the sunpy website (\url{http://sunpy.org}) or can be installed using 
the Python package index (\mbox{\url{http://pypi.python.org/pypi}}).
%schriste - the mbox prevents latex from breaking the url across different lines.

>>>>

\section{Introduction}
Science, including solar physics and astrophysics, is driven by the analysis 
and summarizing of data. Modern advances in sensor technology combined with 
the availability of cheap storage and large bandwidth communications has led to 
a rapid and overwhelming increase in amount of data for scientists to analyse.
For example, NASA's Solar 
Dynamics Observatory (SDO) satellite records over 1 TB of data per day. Analysing this 
mountain of data to enable scientific discovery requires increasingly complex software 
tools. Standardised, easy-to-use,
and transparent software tools are a key enabling technology
so that the scientific community can build upon a common foundation.
%schriste - the opening paragraph still needs work

SunPy aims to provide a free, open-source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breadth of high-quality science-focused packages 
written in that language. Python is a general-purpose, 
powerful and easy-to-learn high-level programming language.
According to the 
2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 Python is one of the top ten most popular programming languages in the world 
and is widely used outside of scientific fields in web development, education 
and `big data' analytics.

The development of a package such as SunPy in Python is made possible by the 
rich ecosystem of scientific packages available in Python. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language: 
\textit{e.g.}, array manipulation, core numerical algorithms and visualisation. 
Built upon these foundations, packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

The design philosophy of SunPy is to provide a clean, simple-to-use, and well 
structured package that provides the \textit{core} tools for solar-physics data analysis. The 
primary focus of SunPy's early development is to provide specialised, linked, 
data types that allow the acquisition, processing and visualisation of all types 
of solar data.

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project, as well as future plans. The latest release of SunPy (0.4)
can be downloaded from
the official website (\url{http://sunpy.org}) or can be installed using 
the Python package index (\mbox{\url{http://pypi.python.org/pypi}}).
%schriste - the mbox prevents latex from breaking the url across different lines.

### 2014-01-28 2-3-Spectra.tex

Spectroscopy consists in the study of the radiative energy related to the wavelength.
An spectrum is normally obtained by observing a range of frequencies at the same time; 
it can be measured for example in radio wavelengths by a frequency sweep with 
a radio receiver or in visible and ultraviolet by the dispersion of the incident light 
through a diffraction grating or prism.
The analysis of the resultant spectrum can provide properties of the plasma observed 
such temperature, density, speed, etc.
Therefore, spectroscopy provides to the solar physicists invaluable information about 
the composition and the physical properties of the Sun.

SunPy aims to provide support to most of solar spectroscopy instrument, however the 
complexity of this kind of datasets makes very challenging to add support to them all 
in a general way.  
Nevertheless, the \texttt{Spectra} object has been designed to be consistent with the 
other data types available in SunPy.
It was implemented as a wrapper around a data array (numpy ndarray)
because its 1D nature (intensity versus frequency).
Using the \texttt{Spectra} object, SunPy also offers a 2D array object called \texttt{Spectrogram},
which it is built as a group of spectrum meassured over time.
These two objects were implemented by funding provided by the Astrophysics Research 
Group at Trinity College Dublin.
As yet, SunPy \texttt{Spectrogram} object supports radio spectra from the e-Callisto 
solar radio spectrometer network(\url{http://www.e-callisto.org/}).
and STEREO/SWAVES spectrograms.
Like for the \texttt{Lightcurve} object, the \texttt{Spectrogram} object has been
built so each instrument initialises from a sub-class containing the instrument-specific 
functionalities while offering common functionalities to them all to help its analysis, 
this includes:
read the data; %(fits format for e-Callisto, ASCII for STEREO/SWAVES); 
join different time ranges and frequencies; 
frequency dependent background subtraction;  
linearise of frequency axis and automated downsample for proper visualisation on a normal computer screen;  
and point-and-click functionality to save pixels of interest for further analysis.
In addition, this object is equipped with a download data interface for e-Callisto that 
only requires the observatory name and the desired time-range. 
STEREO/SWAVES data can be obtained via the virtual solar observatory (see \S\ref{ssec:vso}).

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram} object retrieves
from the online data archive the files that are between the time range specified and
the observatory of interest.  In this case the data is requested from \textit{BIR} which is the code name of the
Rosse Observatory at Birr Castle (Ireland)(\url{http://www.rosseobservatory.ie}).
When the data is requested using the \texttt{.from\_range} function, the object merges
into a single spectrogram all the files downloaded, each with 15 minutes observation data 
and in some cases with different files for different frequency ranges.  
In the example shown here BIR observed in two frequency ranges: 20 - 90\,MHz and 55 - 355\,MHz.
Since the data is not evenly spaced in the frequency range, the \texttt{Spectrogram} object
linearise the frequency axis for a better analisys, as shown in the resultant figures.
It is also shown in the example the effect of the implemented background substraction and
\texttt{.peek} method.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
from SunPy.spectra.sources.callisto import CallistoSpectrogram
tstart = "2011-06-07T06:00:00"
tend = "2011-06-07T07:45:00"
callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
callisto_nobg = callisto.substract_bg()
callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Demonstration of how \texttt{CallistoSpectrogram} object retrieve the data
for the time range and observatory requested, merge it all and removes the background
signal.}
\label{code:spectra}
\end{listing}

>>>>

%ayshih - this whole paragraph can be deleted
Spectroscopy consists in the study of the radiative energy related to the wavelength.
An spectrum is normally obtained by observing a range of frequencies at the same time; 
it can be measured for example in radio wavelengths by a frequency sweep with 
a radio receiver or in visible and ultraviolet by the dispersion of the incident light 
through a diffraction grating or prism.
The analysis of the resultant spectrum can provide properties of the plasma observed 
such temperature, density, speed, etc.
Therefore, spectroscopy provides to the solar physicists invaluable information about 
the composition and the physical properties of the Sun.

SunPy aims to provide broad support for solar spectroscopy instruments, however the 
variety and complexity of such datasets makes it very challenging to add support to them all 
in a general way.
The \texttt{spectra} module implements a \texttt{Spectrum} class for 1D data
(intensity versus frequency) and a \texttt{Spectrogram} class for 2D data
(intensity versus time and frequency.
Each of these classes use a \texttt{numpy.ndarray} class as its \texttt{.data} attribute.
These two classes were implemented by funding provided by the Astrophysics Research 
Group at Trinity College Dublin.

As of SunPy 0.4, the \texttt{Spectrogram} class supports radio spectra from the e-Callisto 
solar radio spectrometer network (\url{http://www.e-callisto.org/})
and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so each instrument initialises using a subclass containing the instrument-specific 
functionalities.
The common functionality provided by the base \texttt{Spectrogram} class includes
reading data,
joining different time ranges and frequencies,
performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram} object retrieves
from the online data archive the files that are between the time range specified and
the observatory of interest.  In this case the data is requested from \textit{BIR},
which is the code name of the
Rosse Observatory at Birr Castle in Ireland (\url{http://www.rosseobservatory.ie}).
When the data is requested using the \texttt{.from\_range} function, the object merges
into a single spectrogram all the files downloaded, each with 15 minutes observation data 
and in some cases with different files for different frequency ranges.  
In the example shown here BIR observed in two frequency ranges: 20--90\,MHz and 55--355\,MHz.
Since the data is not evenly spaced in the frequency range, the \texttt{Spectrogram} object
linearise the frequency axis for a better analisys, as shown in the resultant figures.
It is also shown in the example the effect of the implemented background substraction and
\texttt{.peek} method.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from SunPy.spectra.sources.callisto import CallistoSpectrogram
>>> tstart, tend = "2011-06-07T06:00:00", "2011-06-07T07:45:00"
>>> callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
>>> callisto_nobg = callisto.substract_bg()
>>> callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the data
for the time range and observatory requested, merges it all and removes the background
signal.}
\label{code:spectra}
\end{listing}

### 2014-01-28 3-1-VSO.tex

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from SunPy.net import vso
>>> client = vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 10:30'
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend),
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
40
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type
----------           --------             ------  ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA
...
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronagraphs, path = pathformat)
\end{minted}
\caption{Example of querying a single instrument over a time range and downloading the data}
\label{code:vso_query_simple}
\end{listing}

%schriste - what about just listing the sources as an array to instead of doing a show
%           to make the point that the | query worked appropriately? That would make this
%            listing much shorter.
\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

Querying two wavelength ranges over the same time range
>>> wavelength_condition = ( vso.attrs.Wave(125, 135) |
...                          vso.attrs.Wave(165, 175) ) # in angstroms
>>> euv = client.query(vso.attrs.Time(tstart, tend),
...                    wavelength_condition)
>>> euv.num_records()
3989
>>> euv.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 00:00:00  2011-06-08 00:00:00  SDO       EVE         FULLDISK
...
2011-06-07 05:31:09  2011-06-07 05:31:19  PROBA2    SWAP        FULLSUN
...
2011-06-07 10:30:00  2011-06-07 10:30:01  SDO       AIA         FULLDISK
\end{minted}
\caption{Examples of advanced VSO queries using attribute objects,
continuing from listing~\ref{code:vso_query_simple}.}
\label{code:vso_query_advanced}
\end{listing}

>>>>

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from SunPy.net import vso
>>> client = vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 10:30'
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend),
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
40
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type
----------           --------             ------  ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA
...
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronagraphs, path = pathformat)
\end{minted}
\caption{Example of querying a single instrument over a time range and downloading the data}
\label{code:vso_query_simple}
\end{listing}

%schriste - what about just listing the sources as an array to instead of doing a show
%           to make the point that the | query worked appropriately? That would make this
%            listing much shorter.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
#Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

#Querying two wavelength ranges over the same time range
>>> wavelength_condition = ( vso.attrs.Wave(125, 135) |
...                          vso.attrs.Wave(165, 175) ) # in angstroms
>>> euv = client.query(vso.attrs.Time(tstart, tend),
...                    wavelength_condition)
>>> euv.num_records()
3989
>>> euv.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 00:00:00  2011-06-08 00:00:00  SDO       EVE         FULLDISK
...
2011-06-07 05:31:09  2011-06-07 05:31:19  PROBA2    SWAP        FULLSUN
...
2011-06-07 10:30:00  2011-06-07 10:30:01  SDO       AIA         FULLDISK
\end{minted}
\caption{Examples of advanced VSO queries using attribute objects,
continuing from listing~\ref{code:vso_query_simple}.}
\label{code:vso_query_advanced}
\end{listing}

### 2014-01-28 3-4-Helioviewer.tex

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using the following hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  The full list of image data sources available
can be found through the code snippet, Example
\ref{code:hv:datasources}.
%schriste - what is the point of the following listing? remove?
\begin{listing}[H]
\begin{minted}{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> datasources = hv.get_data_sources()

@@

\begin{listing}[H]
\begin{minted}{python}
hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]", x0=0, y0=0,
 width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{python}
hv.download_png('2099/01/01', 6, 
"[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50], \
[SOHO,LASCO,C2,white-light,1,100]", x0=0, y0=0, width=768, height=768)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00', observatory='SDO',
>>>                            instrument='HMI', detector='HMI', 
>>>                            measurement='continuum')
>>> from sunpy.map import Map
>>> hmi = Map(filepath)
>>> hmi.submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}

>>>>

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using the following hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  The full list of image data sources available
can be found through the code snippet, Example
\ref{code:hv:datasources}.
%schriste - what is the point of the following listing? remove?
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> datasources = hv.get_data_sources()

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> hv.download_png('2099/01/01', 6,
...                 "[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50]"+
...                 "[SOHO,LASCO,C2,white-light,1,100]",
...                 x0=0, y0=0, width=768, height=768)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00', observatory='SDO',
>>>                            instrument='HMI', detector='HMI', 
>>>                            measurement='continuum')
>>> from sunpy.map import Map
>>> hmi = Map(filepath)
>>> hmi.submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}

### 2014-01-28 5-Dev.tex

\section{Development and Community}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub (\url{http://github.com}) and git\footnote{\url{http://git-scm.com/}}
as its distributed version control software.

The continued success of an open source projects depends on many factors of which 
three of the most important are (1) utility and quality of the code, (2) documentation, and (3) an
active community \cite{bangerth2013}. Several tools (some specific to Python) are used by
Sunpy to make achieving these goals more accessible. To maintain high quality code, a 
transparent and collaborative development workflow made possible by Github is used (described
further below). Sunpy includes a number of unit tests which test the functionality 
of the code base. Additionally, Sunpy makes use of `continuous integration' provided by
Travis(\url{http://travis-ci.org}), a process by which the addition of any new code 
automatically triggers a comprehensive review of all unit tests. If any single test
fails then the community is alerted before the code is accepted. The unit test coverage is
also monitored by the \href{http://coveralls.io}{Coveralls} service.

Quality documentation is
one of the most important factors determining the success of any software project. 
Thankfully powerful tools already exist in Python to support this thanks to native
Python's focus on its own documentation. Sunpy makes use of the Sphinx (\url{http://sphinx-doc.org})
documentation generator. Sphinx uses reStructuredText as its markup language, 
an easy-to-read, what-you-see-is-what-you-get plaintext markup syntax. It supports
many output formats most notably HTML as well as PDF, and ePub and provides a rich
hierarchically structured view of in-code documentation strings. The Sunpy documentation 
s built automatically and hosted on Readthedocs (\url{http://readthedocs.org})
at \url{http://docs.sunpy.org}.

In order to maintain an active community communication is key.  The Sunpy community makes
use of a number of channels to communicate. For immediate communications, an active IRC chat
room (\#SunPy) is hosted on freenode.net. For more involved or less immediate needs such as
developer comments or discussion, a open mailing list is used hosted on Google groups. 
Discussion of bugs and bug tracking are provided directly on Github as well as new code
review discussions.

Many of the larger features in SunPy have been developed 
with the support of both the ESA's Summer Of Code In Space (SOCIS) and the Google
Summer of Code (GSOC). Initial development of SunPy's VSO and HEK 
implementations were funded by the SOCIS program as well as a prototyped GUI 
and N-Dimensional datatype implementations. Sunpy was supported by GSOC in 2013 
through the Python Software Foundation (PSF). It supported development of the HELIO and 
database sub-packages.

\subsection{Development Workflow}
The development workflow that GitHub enables is simple and transparent. The 
primary source code repository \url{http://github.com/sunpy/sunpy} can be 
`forked' by anyone with a GitHub account who can then modify their fork as they 
wish. Upon completion of any modifications it is possible for a user to create 
a 'Pull Request' (\url{http://github.com/sunpy/sunpy/pulls}) to the main SunPy 
repository in order to incorporate their modification into the Sunpy code base. 
Pull requests are reviewed by the Sunpy community before being accepted. Community consensus
is required to add new features.

@@

\begin{enumerate}
	\item  The code must follow 	
	\href{http://www.python.org/dev/peps/pep-0008/}{PEP8} the Python style 
	guidelines, this keeps all the code in Sunpy styled in a consistent manner and easy to read.

>>>>

\section{Development and Community}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub (\url{http://github.com}) and Git\footnote{\url{http://git-scm.com/}}
as its distributed version control software.

The continued success of an open-source project depends on many factors;
three of the most important are (1) utility and quality of the code, (2) documentation, and (3) an
active community \cite{bangerth2013}. Several tools (some specific to Python) are used by
SunPy to make achieving these goals more accessible. To maintain high-quality code, a 
transparent and collaborative development workflow made possible by GitHub is used (described
further below).
Additionally, SunPy makes use of `continuous integration' provided by
Travis (\url{http://travis-ci.org}), a process by which the addition of any new code 
automatically triggers a comprehensive review of written unit tests. If any single test
fails then the community is alerted before the code is accepted. The unit test coverage is
also monitored by the \href{http://coveralls.io}{Coveralls} service.

Quality documentation is
one of the most important factors determining the success of any software project. 
Thankfully powerful tools already exist in Python to support this thanks to native
Python's focus on its own documentation. SunPy makes use of the Sphinx (\url{http://sphinx-doc.org})
documentation generator. Sphinx uses reStructuredText as its markup language: 
an easy-to-read, what-you-see-is-what-you-get plaintext markup syntax. It supports
many output formats most notably HTML as well as PDF, and ePub and provides a rich
hierarchically structured view of in-code documentation strings. The SunPy documentation 
is built automatically and hosted on Read the Docs (\url{http://readthedocs.org})
at \url{http://docs.sunpy.org}.

In order to maintain an active community, communication is key.  The SunPy community makes
use of a number of channels to communicate. For immediate communications, an active IRC chat
room (\#SunPy) is hosted on freenode.net. For more involved or less immediate needs, such as
developer comments or discussion, a open mailing list is hosted on Google Groups. 
Discussion of bugs and bug tracking are provided directly on GitHub as well as new
code-review discussions.

Many of the larger features in SunPy have been developed 
with the support of both the ESA's Summer of Code in Space (SOCIS) and the Google
Summer of Code (GSOC). Initial development of SunPy's VSO and HEK 
implementations were funded by the SOCIS program as well as a prototyped GUI 
and N-Dimensional datatype implementations. SunPy was supported by GSOC in 2013 
through the Python Software Foundation (PSF). It supported development of the HELIO and 
database sub-packages.

\subsection{Development Workflow}
The development workflow that GitHub enables is simple and transparent. The 
primary source code repository \url{http://github.com/sunpy/sunpy} can be 
`forked' by anyone with a GitHub account who can then modify their fork as they 
wish. Upon completion of any modifications it is possible for a user to create 
a 'Pull Request' (\url{http://github.com/sunpy/sunpy/pulls}) to the main SunPy 
repository in order to incorporate their modification into the Sunpy code base. 
Pull requests are reviewed by the SunPy community before being accepted. Community consensus
is required to add new features.

@@

\begin{enumerate}
	\item  The code must follow 	
	\href{http://www.python.org/dev/peps/pep-0008/}{PEP8} the Python style 
	guidelines, this keeps all the code in SunPy styled in a consistent manner and easy to read.
